Calculating life expectancy and exploring missingness

### 2016-11-04 Code/1_Explore_Data_v2.Rmd

Add the total population (across age brackets) to the dat2 dataset
```{r}
dat3 <- dat2 %>% ungroup() %>% 
  group_by(StateYearRaceSex) %>%
  summarise(Pop_across_age = sum(Population))

@@

```{r calculate_life_expectancy}
source("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Code/life_expectancy_functions.R")

list.life.tables <- vector(mode = "list", length = length(non.zero.stratums))
list.index <- 1

@@

    list.cod.decomp.tables[[i]] <- cause_of_death_decomp(life.table1 = as.data.frame(list.life.tables[[black.index]]),
                                                         life.table2 = as.data.frame(list.life.tables[[white.index]]), 
                                                         decomp.table = as.data.frame(list.age.decomp.tables[[i]]),
                                                         cod.table = ct.both, 
                                                         age.colname.cod.table = "Age2", 
                                                         COD.colname.cod.table = "COD2", 
                                                         prop1.colname.cod.table = "prop.black", 
                                                         prop2.colname.cod.table = "prop.white")
  } else {
    list.cod.decomp.tables[[i]] <- "Couldn't compare life tables"
  }  
  print(i)
}
```

```{r  update_decomp_for_plotting}
#update the decomposition table for plotting purposes
for(i in 1:dim(paired.ids)[1]) {
    if(list.cod.decomp.tables[[i]][1] != "Couldn't compare life tables") { 
        list.cod.decomp.tables[[i]] <- 
                              make_dataset_cod_plot(cod.decomp.table = list.cod.decomp.tables[[i]], 
                              age.groups = "Ages", 
                              cause.of.death = "Cause.of.death")  
    } 
  print(i)
}  
```

@@

```{r save_workspace3}
save.image(file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/Sep6_BWgap.Rdata")
```

```{r}
save(dat.clean, dat2, BlackWhite, paired.ids, list.age.decomp.tables, list.cod.decomp.tables,
     file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/main_datasets.Rdata")
```

```{r investigate_problem}

list.cod.decomp.tables[[which(paired.ids$StateYearRaceSexBlack == "Washington DC.2013.Black.Male")]]
#index == 2115 fyi

View(dat.clean[dat.clean$State2 == "Washington DC" & dat.clean$Year3 == 2013 & dat.clean$Sex2 == "Male",])
#see the problem is that white males at 1-4 years have no deaths and prop deaths due to each cod == NaN.

list.age.decomp.tables[[which(paired.ids$StateYearRaceSexBlack == "Washington DC.2013.Black.Male")]]

View(list.life.tables[[which(non.zero.stratums == "Washington DC.2013.White.Male")]])

>>>>

Add the total population (across age brackets) to the dat2 dataset.

```{r create_var_pop_across_age}
dat3 <- dat2 %>% ungroup() %>% 
  group_by(StateYearRaceSex) %>%
  summarise(Pop_across_age = sum(Population))

@@

```{r calculate_life_expectancy}
list.life.tables <- vector(mode = "list", length = length(non.zero.stratums))
list.index <- 1

@@

    list.cod.decomp.tables[[i]] <- cause_of_death_decomp(life.table1 = as.data.frame(list.life.tables[[black.index]]),
                                                         life.table2 = as.data.frame(list.life.tables[[white.index]]), 
                                                         decomp.table = as.data.frame(list.age.decomp.tables[[i]]),
                                                         cod.table = ct.both, 
                                                         age.colname.cod.table = "Age2", 
                                                         COD.colname.cod.table = "COD2", 
                                                         prop1.colname.cod.table = "prop.black", 
                                                         prop2.colname.cod.table = "prop.white")
  } else {
    list.cod.decomp.tables[[i]] <- "Couldn't compare life tables"
  }  
  #print(i)
}
```

```{r  update_decomp_for_plotting}
#update the decomposition table for plotting purposes
for(i in 1:dim(paired.ids)[1]) {
    if(list.cod.decomp.tables[[i]][1] != "Couldn't compare life tables") { 
        list.cod.decomp.tables[[i]] <- 
                              make_dataset_cod_plot(cod.decomp.table = list.cod.decomp.tables[[i]], 
                              age.groups = "Ages", 
                              cause.of.death = "Cause.of.death")  
    } 
  #print(i)
}  
```

@@

```{r save_workspace}
save.image(file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/Sep6_BWgap.Rdata")
```

```{r save_specified_files}
save(dat.clean, dat2, BlackWhite, paired.ids, list.age.decomp.tables, list.cod.decomp.tables,
     file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/main_datasets.Rdata")
```
